---
title: "ModernDive Ch11: Tell Your Story with Data"
output: html_notebook
---

```{r}
# Load libraries
library(tidyverse)
library(moderndive)
library(skimr)
library(fivethirtyeight)
library(patchwork)
```

### 11.1 Review

### 11.2 Case study: Seattle house prices
#### 11.2.1 Exploratory data analysis: Part I

```{r}
# Look at raw data
glimpse(house_prices)
```

```{r}
# Skim interesting variables
house_prices %>% 
  select(price, sqft_living, condition) %>% 
  skim()
```

```{r}
# Exploratory vizualisations
(house_prices %>% 
   ggplot(aes(price)) +
   geom_histogram(color = "white")) +
  (house_prices %>% 
     ggplot(aes(sqft_living)) +
     geom_histogram(color = "white")) +
  (house_prices %>% 
     ggplot(aes(condition)) +
     geom_bar())
```
```{r}
# Log10 transform the numeric variables because of long right tail
house_prices <-
  house_prices %>% 
  mutate(log10_price = log10(price),
         log10_size = log10(sqft_living))
house_prices
```

```{r}
# Before and after visualization
((house_prices %>% 
   ggplot(aes(price)) +
   geom_histogram(color = "white")) +
  (house_prices %>% 
   ggplot(aes(log10_price)) +
   geom_histogram(color = "white"))) /
  ((house_prices %>% 
   ggplot(aes(sqft_living)) +
   geom_histogram(color = "white")) +
  (house_prices %>% 
   ggplot(aes(log10_size)) +
   geom_histogram(color = "white")))
  
  
```

#### 11.2.2 Exploratory data analysis: Part II

```{r}
(# Plot interaction model
house_prices %>% 
  ggplot(aes(log10_size, log10_price, color = condition)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = FALSE) +
  theme(legend.position = "none")) +
  
(# Plot parallel slopes model
house_prices %>% 
  ggplot(aes(log10_size, log10_price, color = condition)) +
  geom_point(alpha = 0.05) +
  geom_parallel_slopes(se = FALSE))
  
```

```{r}
# Show faceted version of interaction model
house_prices %>% 
  ggplot(aes(log10_size, log10_price, color = condition)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ condition)
```

#### 11.2.3 Regression modeling

```{r}
# Fit model with interaction
price_interaction <-
  house_prices %>% 
  lm(log10_price ~ log10_size * condition, data = .)

get_regression_table(price_interaction)
```

#### 11.2.4 Making predictions

Say you’re a realtor and someone calls you asking you how much their home will sell for. They tell you that it’s in condition = 5 and is sized 1900 square feet. What do you tell them? Let’s use the interaction model we fit to make predictions!

```{r}
# Use predict to get the predicted log10 price
predicted_log10_price_interaction <-
  price_interaction %>% 
  predict(newdata = tibble(condition = factor(5), log10_size = log10(1900))) %>% 
  unname()
predicted_log10_price_interaction
```

```{r}
# unlog the get the real price
10^predicted_log10_price_interaction
```


```{r}
# (LC11.1) Repeat the regression modeling in Subsection 11.2.3 and the prediction making you just did on the house of condition 5 and size 1900 square feet in Subsection 11.2.4, but using the parallel slopes model you visualized in Figure 11.6. Show that it’s $524,807!

# Fit model with interaction
price_parallel_slopes <-
  house_prices %>% 
  lm(log10_price ~ log10_size + condition, data = .)

get_regression_table(price_parallel_slopes)
```
```{r}
predicted_log10_price_parallel_slopes <-
  price_parallel_slopes %>% 
  predict(newdata = tibble(condition = factor(5), log10_size = log10(1900))) %>% 
  unname()
predicted_log10_price_parallel_slopes
```
```{r}
10^predicted_log10_price_parallel_slopes
```

### 11.3 Case study: Effective data storytelling

#### 11.3.2 US Births in 1999

```{r}
glimpse(US_births_1994_2003)
```

```{r}
US_Biths_1999 <-
  US_births_1994_2003 %>% 
  filter(year == 1999)
```

```{r}
US_Biths_1999 %>% 
  ggplot(aes(date, births)) +
  geom_line() +
  labs(x = "Date", 
       y = "Number of births", 
       title = "US Births in 1999")
```

```{r}
US_Biths_1999 %>% 
  arrange(desc(births))
```

```{r}
# (LC11.2) What date between 1994 and 2003 has the fewest number of births in the US? What story could you tell about why this is the case?

US_births_1994_2003 %>% 
  arrange(births)

# A: Chirstmas day seems to be the day when the least number of births occur.
```



