---
title: "ModernDive Ch6: Multiple Regression"
output: html_notebook
---

```{r}
# Load libraries

library(tidyverse)
library(moderndive)
library(skimr)
library(ISLR)
library(patchwork)

# Load function to clear libraries
source(here::here("clear_libraries.R"))
```

### 6.1 One numerical and one categorical explanatory variable
#### 6.1.1 Exploratory data analysis

```{r}
evals_ch6 <-
  evals %>% 
  select(ID, score, age, gender)
```

```{r}
# Quick look
glimpse(evals_ch6)
```

```{r}
# Random sample
evals_ch6 %>% 
  slice_sample(n = 5)
```

```{r}
# Skim

evals_ch6 %>% 
  select(score, age, gender) %>% 
  skim()
```

```{r}
# Correlation between numerical variables

evals_ch6 %>% 
  get_correlation(formula = score ~ age)

```

```{r}
# Visualize

evals_ch6 %>% 
  ggplot(aes(age, score, color = gender)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Age", y = "Teaching score", color = "Gender")
```
#### 6.1.2 Interaction model
```{r}
# Fit regression model
score_model_interaction <-
  evals_ch6 %>% 
  lm(score ~ age * gender, data = .)

score_model_interaction %>% 
  get_regression_table()
```

#### 6.1.3 Parallel slopes model
```{r}
evals_ch6 %>% 
  ggplot(aes(age, score, color = gender)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE) +
  labs(x = "Age", y = "Teaching score", color = "Gender")
```

```{r}
# Fit linear model

score_model_parallel_slopes <-
  evals_ch6 %>% 
  lm(score ~ age + gender, data = .)

score_model_parallel_slopes %>% 
  get_regression_table()
```

#### 6.1.4 Observed/fitted values and residuals
```{r}
regression_points <- 
  score_model_interaction %>% 
  get_regression_points()

regression_points
```

```{r}
# (LC6.1) Compute the observed values, fitted values, and residuals not for the interaction model as we just did, but rather for the parallel slopes model we saved in score_model_parallel_slopes.

score_model_parallel_slopes %>% 
  get_regression_points()
```

### 6.2 Two numerical explanatory variables
#### 6.2.1 Exploratory data analysis

```{r}
credit_ch6 <-
  Credit %>% 
  as_tibble() %>% 
  select(
    ID, 
    debt = Balance, 
    credit_limit = Limit,
    income = Income,
    credit_rating = Rating,
    age = Age
  )
```

```{r}
glimpse(credit_ch6)
```

```{r}
credit_ch6 %>% slice_sample(n = 5)
```

```{r}
credit_ch6 %>% 
  select(debt, credit_limit, income) %>% 
  skim() 
```

```{r}
credit_ch6 %>% 
  select(debt, credit_limit, income) %>% 
  cor()

# Note collinearity between explanatory variables (credit_limit, income)
```

```{r}
credit_ch6 %>% 
  ggplot(aes(credit_limit, debt)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Credit limit (in $)", 
    y = "Credit card debt (in $)", 
    title = "Debt and credit limit"
  ) +
  
  credit_ch6 %>% 
  ggplot(aes(income, debt)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Income (in $1000)", 
    title = "Debt and income"
  )
```

```{r}
# (LC6.2) Conduct a new exploratory data analysis with the same outcome variable y debt but with credit_rating and age as the new explanatory variables x1 and x2. What can you say about the relationship between a credit card holder’s debt and their credit rating and age?

credit_ch6 %>% 
  select(credit_rating, age) %>% 
  skim()
```

```{r}
credit_ch6 %>% 
  select(debt, credit_rating, age) %>% 
  cor()

# High correlation between debt-credit_rating, low between debt-age
# Low correlation between explanatory variables
```

```{r}
credit_ch6 %>% 
  ggplot(aes(credit_rating, debt)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Credit rating", 
    y = "Credit card debt (in $)", 
    title = "Debt and credit rating"
  ) +
  
  credit_ch6 %>% 
  ggplot(aes(age, debt)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Age", 
    title = "Debt and age"
  )
```

#### 6.2.2 Regression plane
```{r}
# Fit regression model

debt_model <- 
  credit_ch6 %>% 
  lm(debt ~ credit_limit + income, data = .)

debt_model %>% 
  get_regression_table()
```

```{r}
# (LC6.3) Fit a new simple linear regression using lm(debt ~ credit_rating + age, data = credit_ch6) where credit_rating and age are the new numerical explanatory variables x1 and x2. Get information about the “best-fitting” regression plane from the regression table by applying the get_regression_table() function. How do the regression results match up with the results from your previous exploratory data analysis?

credit_ch6 %>% 
  lm(debt ~ credit_rating + age, data = .) %>% 
  get_regression_table()

# Interptretation: When taking into account all other explanatory variables, for every increase in one unit in credit_rating, there is an increase of ~2.6 USD in debt AND when taking into consideration all other explanatory variables, there is a decrease of ~2.4 USD in debt for every added year in age.

# Answer: The slope for age-debt doesn't agree with the exploratory analysis where there was a small positive association between the variables (now clearly negative).
```

#### 6.2.3 Observed/fitted values and residuals

```{r}
debt_model %>% 
  get_regression_points()
```

### 6.3 Related topics
#### 6.3.1 Model selection using visualizations
```{r}
# MA Shools

# interaction model
MA_schools %>% 
  ggplot(aes(perc_disadvan, average_sat_math, color = size)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Percent economically disadvantaged", y = "Math SAT Score", 
       color = "School size", title = "Interaction model") 
```
```{r}
# Parallel slopes model

MA_schools %>% 
   ggplot(aes(perc_disadvan, average_sat_math, color = size)) +
  geom_point(alpha = 0.25) +
  geom_parallel_slopes(se = FALSE) +
  labs(x = "Percent economically disadvantaged", y = "Math SAT Score", 
       color = "School size", title = "Parallel slopes model") 
```

```{r}
# Side-by-side

(# interaction model
MA_schools %>% 
  ggplot(aes(perc_disadvan, average_sat_math, color = size)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Percent economically disadvantaged", y = "Math SAT Score", 
       color = "School size", title = "Interaction model")) +
  (# Parallel slopes model
    MA_schools %>% 
   ggplot(aes(perc_disadvan, average_sat_math, color = size)) +
  geom_point(alpha = 0.25) +
  geom_parallel_slopes(se = FALSE) +
  labs(x = "Percent economically disadvantaged", y = "Math SAT Score", 
       color = "School size", title = "Parallel slopes model"))
```
```{r}
model_2_interaction <-
  MA_schools %>% 
  lm(average_sat_math ~ perc_disadvan * size, data = .)

model_2_interaction %>% 
  get_regression_table()
```

```{r}
model_2_parallel_slopes <-
  MA_schools %>% 
  lm(average_sat_math ~ perc_disadvan + size, data = .)

model_2_parallel_slopes %>% 
  get_regression_table()
```

#### 6.3.2 Model selection using R-squared

```{r}
model_2_interaction %>% 
  get_regression_points()
```

```{r}
# Compute variance for obs, expected, and residual
model_2_interaction %>% 
  get_regression_points() %>% 
  summarise(
    var_y = var(average_sat_math), 
    var_y_hat = var(average_sat_math_hat), 
    var_residual = var(residual)
  )
```

```{r}
# R-squared for interaction model
model_2_interaction %>% 
  get_regression_summaries()
```

```{r}
# R-squared for parallel slopes model
model_2_parallel_slopes%>% 
  get_regression_summaries()
```

#### 6.3.3 Correlation coefficient
```{r}
# Using dollars instead of 1000xdollars doesn't change the correlation
credit_ch6 %>% 
  select(debt, income) %>% 
  mutate(income = income * 1000) %>% 
  cor()
```

```{r}
# The correlation coefficient is invariant to linear transformations
credit_ch6 %>% 
  select(debt, income) %>%  
  cor()
```

#### 6.3.4 Simpson’s Paradox
```{r}
credit_ch6 %>% 
  ggplot(aes(credit_limit)) +
  geom_histogram(color = "white") 
```

```{r}
credit_ch6 %>% 
  ggplot(aes(income, debt))+ 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Credit card debt (in $)",
    y = "Income (in $1000",
  ) +

credit_ch6 %>% 
  mutate(
    credit_limit_bracket = case_when(
      credit_limit < 3088 ~ "low",
      credit_limit < 4622 ~ "medium-low",
      credit_limit < 5873 ~ "medium-high",
      TRUE ~ "high"
    )
  ) %>% 
  ggplot(aes(income, debt, color = credit_limit_bracket)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Credit card debt (in $)",
    y = "Income (in $1000",
    color = "Credit limit bracket"
  )
```

```{r}
clear_libraries()
```
