---
title: "ModernDive Ch8: Bootstrapping and Confidence Intervals"
output: html_notebook
---

```{r}
# Load libraries
library(tidyverse)
library(moderndive)
library(infer)
library(patchwork)

# Load function to clear libraries
source(here::here("clear_libraries.R"))
```

### 8.1 Pennies activity
#### 8.1.1 What is the average year on US pennies in 2019?

```{r}
pennies_sample
```

```{r}
# Exploratory visualization

pennies_sample %>% 
  ggplot(aes(year)) +
  geom_histogram(binwidth = 10, color = "white")
```

```{r}
# Mean year
x_bar <-
  pennies_sample %>% 
  summarise(mean_year = mean(year))
x_bar
```

#### 8.1.2 Resampling once
```{r}
pennies_resample <-
  tibble(
    year = c(
      1976,
      1962,
      1976,
      1983,
      2017,
      2015,
      2015,
      1962,
      2016,
      1976,
      2006,
      1997,
      1988,
      2015,
      2015,
      1988,
      2016,
      1978,
      1979,
      1997,
      1974,
      2013,
      1978,
      2015,
      2008,
      1982,
      1986,
      1979,
      1981,
      2004,
      2000,
      1995,
      1999,
      2006,
      1979,
      2015,
      1979,
      1998,
      1981,
      2015,
      2000,
      1999,
      1988,
      2017,
      1992,
      1997,
      1990,
      1988,
      2006,
      2000
    )
  )
```

```{r}
pennies_resample %>% 
  ggplot(aes(year)) +
  geom_histogram(binwidth = 10, color = "white") +
  labs(title = "Resample of 50 pennies") +
  # This doesn't reproduce the figure from the book

pennies_sample %>% 
  ggplot(aes(year)) +
  geom_histogram(binwidth = 10, color = "white") +
  labs(title = "Original sample of 50 pennies")
```

```{r}
# Resample mean
pennies_resample %>% 
  summarise(mean_year = mean(year))

# This is not the same result as in the book
```

#### 8.1.3 Resampling 35 times

```{r}
resampled_means <-
  pennies_resamples %>% 
  group_by(name) %>% 
  summarise(mean_year = mean(year))
resampled_means
```

```{r}
# Visualize resampling (bootstrap) distribution
resampled_means %>% 
  ggplot(aes(mean_year)) +
  geom_histogram(binwidth = 1, boundary = 1990, color = "white") +
  labs(x = "Sampled mean year")
```

#### 8.1.4 What did we just do?

### 8.2 Computer simulation of resampling
#### 8.2.1 Virtually resampling once

```{r}
virtual_resample <-
  pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE)
virtual_resample
```

```{r}
virtual_resample %>% 
  summarise(resample_mean = mean(year))
```

#### 8.2.2 Virtually resampling 35 times
```{r}
# Create virtual resamples
virtual_resamples <-
  pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE, reps = 35)
virtual_resamples
```

```{r}
# Count means for each replicate
virtual_resampled_means <-
  virtual_resamples %>% 
  group_by(replicate) %>% 
  summarise(mean_year = mean(year))
virtual_resampled_means
```

```{r}
# Visualize bootstrap distribution
virtual_resampled_means %>% 
  ggplot(aes(mean_year)) +
  geom_histogram(binwidth = 1, boundary = 1990, color = "white") +
  labs(x = "Resample mean year")
```

#### 8.2.3 Virtually resampling 1000 times

```{r}
# Repeat resampling 1000 times
virtual_resamples <-
  pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE, reps = 1000)
virtual_resamples
```

```{r}
# Compute 1000 sample means
virtual_resampled_means <-
  virtual_resamples %>% 
  group_by(replicate) %>% 
  summarise(mean_year = mean(year))
virtual_resampled_means
```

```{r}
# Combined to one pipeline
virtual_resampled_means <-
  pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE, reps = 1000) %>% 
  group_by(replicate) %>% 
  summarise(mean_year = mean(year))
virtual_resampled_means
```

```{r}
# Visualize bootstrap distibution from the 1000 replicates
virtual_resampled_means %>% 
  ggplot(aes(mean_year)) +
  geom_histogram(binwidth = 1, boundary = 1990, color = "white") +
  labs(x = "sample mean")
```

```{r}
virtual_resampled_means %>% 
  summarise(mean_of_means = mean(mean_year))
```

```{r}
# (LC8.1) What is the chief difference between a bootstrap distribution and a sampling distribution?

# A: A Sampling distribution is derived from individual samples all drawn from the actual population. A Bootstrap distribution is derived from one sample drawn from the population that is then sampled with replacement to get an approximation of the sampling distribution

# (LC8.2) Looking at the bootstrap distribution for the sample mean in Figure 8.14, between what two values would you say most values lie?

# A: 1991-2000
```

### 8.3 Understanding confidence intervals
#### 8.3.1 Percentile method

#### 8.3.2 Standard error method
```{r}
# Compute SE of bootstrap distribution
virtual_resampled_means %>% 
  summarise(SE = sd(mean_year))
```

```{r}
# (LC8.3) What condition about the bootstrap distribution must be met for us to be able to construct confidence intervals using the standard error method?

# A: The bootstrap distribution must approximate a normal distribution

# (LC8.4) Say we wanted to construct a 68% confidence interval instead of a 95% confidence interval for μ. Describe what changes are needed to make this happen. Hint: we suggest you look at Appendix A.2 on the normal distribution.

# A: ~68% of the values will be inside +- 1SD in a normal distribution, so we would use mean +- SE as our confidence interval
```

### 8.4 Constructing confidence intervals
#### 8.4.1 Original workflow
#### 8.4.2 infer package workflow

```{r}
# Using dplyr
pennies_sample %>% 
  summarise(stat = mean(year))
```

```{r}
# Using infer
pennies_sample %>% 
  specify(response = year) %>% 
  calculate(stat = "mean")
```

```{r}
# 1. SPECIFY - variables
pennies_sample %>% 
  specify(response = year) # or (response = yera ~ NULL)
```

```{r}
# 2. GENERATE - data/replicates
pennies_sample %>% 
  specify(response = year) %>% 
  generate(reps = 1000, type = "bootstrap")
```

```{r}
# Compare to original workflow
pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE, reps = 1000)
```

```{r}
# 3. CALCULATE - summary statistics
bootstrap_distribution <-
  pennies_sample %>% 
  specify(response = year) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")
bootstrap_distribution
```

```{r}
# Compare to original workflow
pennies_sample %>% 
  rep_slice_sample(n = 50, replace = TRUE, reps = 1000) %>% 
  group_by(replicate) %>% 
  summarise(stat = mean(year))

```

```{r}
# 4. VISUALIZE - the results
  bootstrap_distribution %>% 
  visualize() 
```

```{r}
# Compare to original workflow
bootstrap_distribution %>% 
  ggplot(aes(stat)) +
  geom_histogram(binwidth = 1, color = "white")
```

#### 8.4.3 Percentile method with infer

```{r}
# 95% CI with the percentile method
percentile_ci <-
  bootstrap_distribution %>% 
  get_confidence_interval(level = 0.95, type = "percentile")
percentile_ci
```

```{r}
# Visualize 95% CI (using percentile_ci)
bootstrap_distribution %>% 
  visualize() +
  shade_confidence_interval(endpoints = percentile_ci) 
```

```{r}
# Shorter version (shade_ci)
bootstrap_distribution %>%
  visualize() +
  shade_ci(
    endpoints = percentile_ci,
    color = "hotpink",
    fill = "khaki",
    size = 1,
    alpha = 0.5
  )
```

```{r}
# 95% CI with the SE method
standard_error_ci <-
  bootstrap_distribution %>% 
  get_ci(type = "se", point_estimate = x_bar) # Note abbreviated function name
standard_error_ci
```

```{r}
# Visualize 95% CI using the SE method
bootstrap_distribution %>% 
  visualize() +
  shade_ci(endpoints = standard_error_ci)
```

```{r}
# (LC8.5) Construct a 95% confidence interval for the median year of minting of all US pennies. Use the percentile method and, if appropriate, then use the standard-error method.

# Note: Book says SE method is not possible b/c the distribution is not bell-shaped

bootstrap_distribution_median <-
  pennies_sample %>%
  specify(response = year) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "median")

ci_95_percentile <-
  bootstrap_distribution_median %>% 
  get_ci(level = 0.95, type = "percentile")

ci_95_percentile
```

```{r}
bootstrap_distribution_median %>%
  visualize() +
  shade_ci(endpoints = ci_95_percentile)
```

### 8.5 Interpreting confidence intervals

```{r}
# Revisit the bowl sampling from Ch7; find true pop. proportion
bowl %>% 
  summarise(p_red = mean(color == "red"))
```

#### 8.5.1 Did the net capture the fish?

```{r}
# Ilyas and Yohan’s sample
bowl_sample_1
```

```{r}
# Specify
bowl_sample_1 %>% 
  specify(response = color, success = "red")
```

```{r}
# Generate
bowl_sample_1 %>% 
  specify(response = color, success = "red") %>% 
  generate(reps = 1000, type = "bootstrap")

```

```{r}
# Calculate
sample_1_bootstrap <-
  bowl_sample_1 %>% 
  specify(response = color, success = "red") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")
sample_1_bootstrap
```

```{r}
# Get 95% CI
percentile_ci_1 <-
  sample_1_bootstrap %>% 
  get_ci(level = 0.95, type = "percentile")
percentile_ci_1

# This CI contains the true population parameter, 0.375
```

```{r}
# Visualize

sample_1_bootstrap %>% 
  visualize(bins = 15) +
  shade_ci(endpoints = percentile_ci_1) +
  geom_vline(xintercept = 0.42, linetype = "dashed")
```

```{r}
# Will every 95% confidence interval capture the true pop. parameter
bowl_sample_2 <-
  bowl %>% 
  rep_slice_sample(n = 50)
bowl_sample_2
```

```{r}
sample_2_bootstrap <-
  bowl_sample_2 %>% 
  specify(response = color, success = "red") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")
sample_2_bootstrap
```

```{r}
# The real population proportion is again within the CI
percentile_ci_2 <-
  sample_2_bootstrap %>% 
  get_ci(level = 0.95, type = "percentile")
percentile_ci_2
```

#### 8.5.2 Precise and shorthand interpretation
```{r}
# Precise interpretation: If we repeated our sampling procedure a large number of times, we expect about 95% of the resulting confidence intervals to capture the value of the population parameter.

# incorrect interpretation is: There is a 95% probability that the confidence interval contains p.

# Short-hand interpretation: We are 95% “confident” that a 95% confidence interval captures the value of the population parameter.
```

#### 8.5.3 Width of confidence intervals

### 8.6 Case study: Is yawning contagious?
#### 8.6.1 Mythbusters study data

```{r}
mythbusters_yawn
```

```{r}
mythbusters_yawn %>% count(group, yawn)
```

#### 8.6.2 Sampling scenario
#### 8.6.3 Constructing the confidence interval

```{r}
# Specify
mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes")
```

```{r}
# Generate
mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  generate(reps = 1000, type = "bootstrap")
```

```{r}
# Calculate
bootstrap_distribution_yawning <-
  mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "diff in props", order = c("seed", "control"))
bootstrap_distribution_yawning 
```

```{r}
# Visualize
bootstrap_distribution_yawning %>% 
  visualize() +
  geom_vline(xintercept = 0)
```

```{r}
# Compute 95% CI using the percentile method
myth_ci_percentile <-
  bootstrap_distribution_yawning %>% 
  get_ci(level = 0.95, type = "percentile")
myth_ci_percentile
```

```{r}
# Compute 95% CI using the SE method
obs_diff_in_props <- mythbusters_yawn %>% 
  specify(formula = yawn ~ group, success = "yes") %>% 
  # generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in props", order = c("seed", "control"))

myth_ci_se <-
  bootstrap_distribution_yawning %>% 
  get_ci(type = "se", point_estimate = obs_diff_in_props)
```

```{r}
# Visualize
bootstrap_distribution_yawning %>% 
  visualize() +
  shade_ci(endpoints = myth_ci_percentile, color = "black") +
  shade_ci(endpoints = myth_ci_se, color = "gray")
```

#### 8.6.4 Interpreting the confidence interval

### 8.7 Conclusion
#### 8.7.1 Comparing bootstrap and sampling distributions

```{r}
# Take 1000 virtual samples of size 50 from the bowl:
virtual_samples <- 
  bowl %>% 
  rep_slice_sample(n = 50, reps = 1000)

# Compute the sampling distribution of 1000 values of p-hat
sampling_distribution <- 
  virtual_samples %>% 
  group_by(replicate) %>% 
  summarize(red = sum(color == "red")) %>% 
  mutate(prop_red = red / 50)

# Visualize sampling distribution of p-hat
sampling_distribution %>% 
  ggplot(aes(prop_red)) +
  geom_histogram(binwidth = 0.05, boundary = 0.4, color = "white") +
  labs(x = "Proportion of 50 balls that were red", 
       title = "Sampling distribution")
```

```{r}
sampling_distribution %>% 
  summarise(se = sd(prop_red))
```

Bootstrap distribution
```{r}
bootstrap_distribution  <-
  bowl_sample_1 %>% 
  specify(response = color, success = "red") %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "prop")
bootstrap_distribution
```

```{r}
bootstrap_distribution %>% 
  visualize()
```

```{r}
bootstrap_distribution %>% 
  summarise(se = sd(stat))
```

Confidence intervals based on 33 tactile samples
```{r}
conf_ints <- 
  tactile_prop_red %>% 
  rename(p_hat = prop_red) %>% 
  mutate(
    n = 50,
    SE = sqrt(p_hat * (1 - p_hat) / n),
    MoE = 1.96 * SE,
    lower_ci = p_hat - MoE,
    upper_ci = p_hat + MoE
  )
conf_ints
```
```{r}
# Did the net capture the fish 95% of the time
conf_ints <-
  conf_ints %>% 
  mutate(captured_pop_prop = lower_ci <= 0.375 & 0.375 <= upper_ci)
mean(conf_ints$captured_pop_prop)

# Almost
```

```{r}
clear_libraries()
```